---
title: "Hawaiʻi Sugar Abandonment Sequence"
author: "D. Nākoa Farrant"
date: '2022-06-29'
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

```{r}
library(plyr)
library(tidyverse) #for datafrmae manipulation
library(data.table) # for fread()
library(Hmisc)
library(corrplot)
library(dplyr)
library(sf)
library(lwgeom)
library(stargazer)
library(kableExtra)
library(viridis) # visualization package for color
library(RColorBrewer)
library(ggspatial) # required for plot annotations like scale bar
library(raster)
library(sp)
library(tmap) # alternative to leaflet to do some interactive maps and maybe select polygons interactively 
library(leaflet) # ideal for manual selection of polygons, but unsure if will have the opportunity to do that 
library(maptools) # for zonal statistics
library(rnaturalearth)
library(rnaturalearthdata)
```

```{r}
HI_coastline <- st_read("coast_n83/coast_n83.shp") %>% 
  st_transform(st_crs(4326)) %>% 
  st_set_precision(1e6) %>% 
  st_make_valid() # class sf and data.frame
```

Goal: Write a function that produces a sequence of erased shapefiles and appropriately labels them

1) Ready in all of the relevant shapefiles
2) Ensure that all shapefiles are in the same projection
3) Sequentially erase all of the shapefiles 
4) Save each shapefile from each point in time


Read in 1900
Erase 1920
Erase 1937
Erase 1980 - do we remove the areas that may or man not be agricultural land in 1980?
Erase 2015

```{r}
s1900 <- st_read("sugar_gee/sugar1900.shp") %>% 
  st_transform(st_crs(4326)) %>% 
  st_set_precision(1e6) %>% 
  st_make_valid() %>% 
  dplyr::select(Year)
s1920 <- st_read("sugar_gee/sugar1920.shp") %>% 
  st_transform(st_crs(4326)) %>% 
  st_set_precision(1e6) %>% 
  st_make_valid() %>% 
  dplyr::select(Year)
s1937 <- st_read("sugar_gee/sugar1937.shp") %>% 
  st_transform(st_crs(4326)) %>% 
  st_set_precision(1e6) %>% 
  st_make_valid() %>% 
  dplyr::select(Year)
ag1980 <- st_read("sugar_gee/alum.shp") %>% 
  st_transform(st_crs(4326)) %>% 
  st_set_precision(1e6) %>% 
  st_make_valid() %>% 
  mutate(Year = 1980)
ag2015 <- st_read("sugar_gee/ag2015.shp") %>% 
  st_transform(st_crs(4326)) %>% 
  st_set_precision(1e6) %>% 
  st_make_valid() %>% 
  mutate(Year = 2015)
ag2020 <- st_read("sugar_gee/ag2020.shp") %>% 
  st_transform(st_crs(4326)) %>% 
  st_set_precision(1e6) %>% 
  st_make_valid() %>% 
  mutate(Year = 2020)
```

```{r}
# Specifically filter for sugar in more recent agricultural shape files
s1980 <- ag1980 %>% 
  filter(commodity == 'S') %>% 
  dplyr::select(Year)

s2015 <- ag2015 %>% 
  filter(CropCatego == "Sugar") %>% 
  dplyr::select(Year)

s2020 <- ag2020 %>% 
  filter(Crops_2020 == 'Sugar') %>% 
  dplyr::select(Year)
```

```{r}
# this binds all of the different sugar shapefiles quickly. Unfortunately some of shapefiles overlap
all_sugar <- do.call("rbind", list(s1900, s1920, s1937, s1980, s2015, s2020))
#st_write(all_sugar, "output_files/all_sugar.gpkg", driver = "gpkg")
```

```{r}
hi_box_poly = st_polygon(
   list(
     cbind(
       c(-160.5, -154.2, -154.2, -160.5, -160.5), 
       c(22.8, 22.8, 18.5, 18.5, 22.8))
     )
) 

hi_box_projected = st_sfc(hi_box_poly, crs = 4326)
```

```{r}
all_sugar$HI <- ifelse(sf::st_intersects(all_sugar, hi_box_projected, sparse = F), 
                                 "Yes", 
                                 "No")
```


```{r}

all_sugar_summarise <- all_sugar %>% 
  group_by(HI) %>% 
  summarise()
```
```{r}
all_sugar_year <- all_sugar %>% 
  mutate(Year = factor(Year, levels = c(2020, 2015, 1980, 1937, 1920, 1900))) %>% 
  group_by(Year) %>% 
  summarize()
```

```{r}
ggplot() +
  geom_sf(data = HI_coastline, fill = "grey60") +
  geom_sf(data = all_sugar_year, aes(color = Year, fill = Year)) +
  scale_color_brewer(palette = "YlOrRd") +
  scale_fill_brewer(palette = "YlOrRd") +
  theme_bw()
```

```{r}
###world
world <- ne_countries(scale = "medium", returnclass = "sf") %>% 
  st_transform(st_crs(4326))
```

```{r}
ggplot() +
  geom_sf(data = world) + 
  geom_sf(data = all_sugar_year, aes(color = Year, fill = Year)) +
  coord_sf(xlim = c(-160.5, -154.2), ylim = c(18.5, 22.8), expand = F) +
  scale_color_brewer(palette = "YlOrRd") +
  scale_fill_brewer(palette = "YlOrRd") +
  labs(x = "Longitude (º)", y = "Latitude (º)") +
  theme_bw()+
  theme(panel.border = element_rect(colour = "black", fill=NA, size=1),
  panel.background = element_rect(fill = "#2A788EFF"), panel.grid.major = element_line(size = 0.5, linetype = 'solid',
                                colour = "#2A788EFF"))+
  annotation_scale(location = "br")
```

```{r}
st_write(all_sugar, "output_files/all_sugar_HI.gpkg", driver = "gpkg")
st_write(all_sugar, "output_files/all_sugar_HI.shp")
```

```{r}
pi1900 <- st_read("pineapple_ag/pineapple_1900.shp")  %>% 
  st_transform(st_crs(4326)) %>% 
  st_set_precision(1e6) %>% 
  st_make_valid() %>% 
  mutate(Year = 1900)
pi1920 <- st_read("pineapple_ag/pineapple_1920.shp")  %>% 
  st_transform(st_crs(4326)) %>% 
  st_set_precision(1e6) %>% 
  st_make_valid() %>% 
  mutate(Year = 1920)
pi1937 <- st_read("pineapple_ag/pineapple_1937.shp") %>% 
  st_transform(st_crs(4326)) %>% 
  st_set_precision(1e6) %>% 
  st_make_valid() %>% 
  mutate(Year = 1937)
ag1980 <- st_read("hi_ag1980/alum.shp")  %>% 
  st_transform(st_crs(4326)) %>% 
  st_set_precision(1e6) %>% 
  st_make_valid() %>% 
  mutate(Year = 1980)
ag2015 <- st_read("hi_ag2015/2015AgBaseline.shp") %>% 
  st_transform(st_crs(4326)) %>% 
  st_set_precision(1e6) %>% 
  st_make_valid() %>% 
  mutate(Year = 2015)
ag2020 <- st_read("HI_Ag_2020/aglanduse_2020.shp") %>% 
  st_transform(st_crs(4326)) %>% 
  st_set_precision(1e6) %>% 
  st_make_valid() %>% 
  mutate(Year = 2020)
```


```{r}
pi1980 <- ag1980 %>% 
  filter(commodity == 'P') %>% 
  dplyr::select(Year)

pi2015 <- ag2015 %>% 
  filter(CropCatego == "Pineapple") %>% 
  dplyr::select(Year)

pi2020 <- ag2020 %>% 
  filter(Crops_2020 == "Pineapple") %>% 
  dplyr::select(Year)
```

Want to identify when a field started being cultivated 

```{r}
#a helper function that erases all y from x
st_erase = function(x, y) {
st_difference(
  st_geometry(x) %>% st_buffer(0),
  st_union(st_combine(st_geometry(y))) %>% st_buffer(0)
)
}
```


```{r}
# Question about st_union here is if all geometries will be valid and are basically "dissolved" boundaries of polygons that exist within each other 
pine1900 <- pi1900 %>% 
  mutate(survey_yr = 1900, earliest_cult = 1891)

pine1920 <- st_erase(pi1920, pi1900) %>%
  st_as_sf() %>% 
  st_transform(st_crs("+proj=utm +ellps=GRS80 +datum=WGS84")) %>% 
  mutate(survey_yr = 1920, earliest_cult = 1901)

pi_pre1937 <- st_union(pi1900, pine1920)

pine1937 <- st_erase(pi1937, pi_pre1937) %>%
  st_as_sf() %>% 
  st_transform(st_crs("+proj=utm +ellps=GRS80 +datum=WGS84")) %>% 
  mutate(survey_yr = 1937, earliest_cult = 1921)

pi_pre1980 <- st_union(pi1900, pine1920, pine1937)

pine1980 <- st_erase(pi1980, pi_pre1980) %>% 
  st_as_sf() %>% 
  st_transform(st_crs("+proj=utm +ellps=GRS80 +datum=WGS84")) %>% 
  mutate(survey_yr = 1980, earliest_cult = 1938)

pi_pre2015 <- st_union(pi1900, pine1920, pine1937, pine1980)

pine2015 <- st_erase(pi2015, pi_pre2015) %>%
  st_as_sf() %>% 
  st_transform(st_crs("+proj=utm +ellps=GRS80 +datum=WGS84")) %>% 
  mutate(survey_yr = 2015, earliest_cult = 1981)

pi_pre2020 <- st_union(pi1900, pine1920, pine1937, pine1980, pine2020)

pine2020 <- st_erase(pi2020, pi_pre2020) %>%
  st_as_sf() %>% 
  st_transform(st_crs("+proj=utm +ellps=GRS80 +datum=WGS84")) %>% 
  mutate(survey_yr = 2020, earliest_cult = 2016)
```

```{r}
pine1920sf <- st_as_sf(pine1920) %>% 
  mutate(survey_yr == 1920)
```



```{r}
ggplot() +
  geom_sf(data = pi1900, fill = "blue", color = "blue") +
  geom_sf(data = pine1920, fill = "green", color = "green") +
  geom_sf(data = pine1937, fill = "red", color = "red") +
  geom_sf(data = pine1980, fill = "yellow", color = "yellow") +
  geom_sf(data = pine2015, fill = "purple", color = "purple") +
  geom_sf(data = pine2020, fill = "pink", color = "pink") +
  geom_sf(data = HI_coastline, color = "black", alpha = 0.1)
```

Potentially useless to do the 2020 erase because there was no change in pineapple area from 2015 other than very minor changes in pineapple boundary. Could be worth seeing if the 2020 ag survey gets updated after March 2022





```{r}
# Pineapple cultivated 1900
pi1900_ab1920 <- st_erase(pi1900, pi1920) %>%
   st_as_sf() %>%
   mutate(start_survey_yr == 1900, earliest_cult == 1891, aban_survey_yr == 1920)

pi1900_ab1937 <- st_erase(pi1900, pi1937) %>% 
  st_erase(pi1900_ab1920) %>% 
  st_as_sf() %>% 
  mutate(start_survey_yr == 1900, earliest_cult == 1891, aban_survey_yr == 1937)

pi1900_ab1980 <- st_erase(pi1900, ag1980) %>% # should it be ag 1980 or pine1980? feels more robust to use ag1980
    st_erase(pi1900_ab1920) %>%
    st_erase(pi1900_ab1937) %>%
    st_as_sf() %>% 
    mutate(start_survey_yr == 1900, earliest_cult == 1891, aban_survey_yr == 1980)
 
pi1900_ab2015 <- st_erase(pine1900, ag2015) %>% 
    st_erase(pi1900_ab1920) %>%
    st_erase(pi1900_ab1937) %>% 
    st_erase(pi1900_ab1980) %>% 
    st_as_sf() %>% 
    mutate(start_survey_yr == 1900, earliest_cult == 1891, aban_survey_yr == 2015)
 
# pi1900_ab2020 <- st_erase(pine1900, ag2020) %>% 
#     st_erase(pi1900_ab1920) %>%
#     st_erase(pi1900_ab1937) %>% 
#     st_erase(pi1900_ab1980) %>% 
#     st_erase(pi1900_ab2015) %>% 
#     st_as_sf() %>% 
#     mutate(start_survey_yr == 1900, earliest_cult == 1891, aban_survey_yr == 2020)


# Pineapple cultivated 1920
pi1920_ab1937 <- st_erase(pi1920, pi1937) %>% 
    st_as_sf() %>% 
    mutate(start_survey_yr == 1920, earliest_cult == 1901, aban_survey_yr == 1937)

pi1920_ab1980 <- st_erase(pi1920, ag1980) %>% # should it be ag 1980 or pine1980? feels more robust to use ag1980
    st_erase(pi1920_ab1937) %>%
    st_as_sf() %>% 
    mutate(start_survey_yr == 1920, earliest_cult == 1901, aban_survey_yr == 1980)
 
pi1920_ab2015 <- st_erase(pine1920, ag2015) %>% 
    st_erase(pi1920_ab1937) %>% 
    st_erase(pi1920_ab1980) %>% 
    st_as_sf() %>% 
    mutate(start_survey_yr == 1920, earliest_cult == 1901, aban_survey_yr == 2015)
 
# pi1920_ab2020 <- st_erase(pine1920, ag2020) %>% 
#     st_erase(pi1900_ab1920) %>%
#     st_erase(pi1900_ab1937) %>% 
#     st_erase(pi1900_ab1980) %>% 
#     st_erase(pi1900_ab2015) %>% 
#     st_as_sf() %>% 
#     mutate(start_survey_yr == 1920, earliest_cult == 1901, aban_survey_yr == 2020)


# Pineapple cultivated 1937
pi1937_ab1980 <- st_erase(pi1937, ag1980) %>% # should it be ag 1980 or pine1980? feels more robust to use ag1980
    st_as_sf() %>% 
    mutate(start_survey_yr == 1937, earliest_cult == 1921, aban_survey_yr == 1980)
 
pi1937_ab2015 <- st_erase(pine1937, ag2015) %>% 
    st_erase(pi1937_ab1980) %>% 
    st_as_sf() %>% 
    mutate(start_survey_yr == 1937, earliest_cult == 1921, aban_survey_yr == 2015)
 
# pi1937_ab2020 <- st_erase(pine1937, ag2020) %>% 
#     st_erase(pi1937_ab1980) %>% 
#     st_erase(pi1937_ab2015) %>% 
#     st_as_sf() %>% 
#     mutate(start_survey_yr == 1920, earliest_cult == 1921, aban_survey_yr == 2020)

# Pineapple cultivated 1980
pi1980_ab2015 <- st_erase(pine1980, ag2015) %>% 
    st_as_sf() %>% 
    mutate(start_survey_yr == 1980, earliest_cult == 1938, aban_survey_yr == 2015)
 
# pi1980_ab2020 <- st_erase(pine1980, ag2020) %>% 
#     st_erase(pi1980_ab2015) %>% 
#     st_as_sf() %>% 
#     mutate(start_survey_yr == 1980, earliest_cult == 1938, aban_survey_yr == 2020)

# Pineapple cultivated 2015
pi2015_ab2020 <- st_erase(pine2015, ag2020) %>% 
    st_as_sf() %>% 
    mutate(start_survey_yr == 2015, earliest_cult == 1981, aban_survey_yr == 2020)
```


```{r}
valid1900_1920 <- st_is_valid(pi1900_ab1920)
valid1900_1937 <- st_is_valid(pi1900_ab1937)
valid1900_1980 <- st_is_valid(pi1900_ab1980)
valid1900_2015 <- st_is_valid(pi1900_ab2015)

valid1920_1937 <- st_is_valid(pi1920_ab1937)
valid1920_1980 <- st_is_valid(pi1920_ab1980)
valid1920_2015 <- st_is_valid(pi1920_ab2015)

valid1937_1980 <- st_is_valid(pi1937_ab1980)
valid1937_2015 <- st_is_valid(pi1937_ab2015)

valid1980_2015 <- st_is_valid(pi1980_ab2015)
```

```{r}
# identify which index/indices in the list have false for st_is_valid
# checked the validity of all of the sf objects and it seems fine
which(valid1937_1980 == FALSE, arr.ind = T)
```


```{r}
ggplot() +
  geom_sf(data = pi1900_ab1920, color = "blue", fill = "blue", alpha = 0.5) +
  geom_sf(data = pi1900_ab1937, color = "red", fill = "red", alpha = 0.5) +
  geom_sf(data = pi1900_ab1980, color = "green", fill = "green", alpha = 0.5) +
  geom_sf(data = pi1900_ab2015, color = "orange", fill = "orange", alpha = 0.5)
```


```{r}
# pine_allaban <- rbind(pi1900_ab1920, pi1900_ab1937, pi1900_ab1980, pi1900_ab2015, pi1900_ab2020, pi1920_ab1937, pi1920_ab1980, pi1920_ab2015, pi1920_ab2020, pi1937_ab1980, pi1937_ab2015, pi1937_ab2020, pi1980_ab2015, pi1980_ab2020, pi2015_ab2020) %>% 
#   mutate(aban_survey_yr = as.factor(aban_survey_yr), earliest_cult = as.factor(earliest_cult), start_survey_yr = as.factor(start_survey_yr))

# run a version without the 2020 erasures that seem to make limited difference
pine_allaban <- rbind(pi1900_ab1920, pi1900_ab1937, pi1900_ab1980, pi1900_ab2015, pi1920_ab1937, pi1920_ab1980, pi1920_ab2015, pi1937_ab1980, pi1937_ab2015, pi1980_ab2015) %>% 
  mutate(aban_survey_yr = as.factor(aban_survey_yr), earliest_cult = as.factor(earliest_cult), start_survey_yr = as.factor(start_survey_yr))
```


```{r}
hi_habstatus <- st_read("cah_habitat_status_poly.shp") %>% 
  st_transform(st_crs("+proj=utm +ellps=GRS80 +datum=WGS84")) %>% 
  st_set_precision(1000000) %>% 
  st_make_valid() # class sf and data.frame
```

```{r}
pine_aban_habstatus <- st_join(pine_allaban, hi_habstatus)
```

```{r}
detach(package:plyr)
```

```{r}
pine_aban_habstatus$VegCover = ifelse(pine_aban_habstatus$Descriptio == "Native Dominated", "Native", 
                                                         ifelse(pine_aban_habstatus$Descriptio == "Native / Alien Mix", "Invasive", "Bare"))

pine_aban_habstatus$aban_yr_num = ifelse(pine_aban_habstatus$aban_survey_yr == "1920", 1920, 
                                                         ifelse(pine_aban_habstatus$aban_survey_yr == "1937", 1937,
                                                                ifelse(pine_aban_habstatus$aban_survey_yr == "1980", 1980, 2015)))

pine_aban_habstatus$earliest_yr_num = ifelse(pine_aban_habstatus$earliest_cult == "1891", 1891, 
                                                         ifelse(pine_aban_habstatus$earliest_cult == "1901", 1901,
                                                                ifelse(pine_aban_habstatus$earliest_cult == "1921", 1921, 1938)))


# calculate time since abandonment relative to 2020
pine_aban_habstatus$TimeSinceAb = 2020 - pine_aban_habstatus$aban_yr_num
pine_aban_habstatus$MaxCultTime = pine_aban_habstatus$aban_yr_num - pine_aban_habstatus$earliest_yr_num
```

Possible approaches to lable specific plantations with a plantation name to add information about the specific closure year
1) Manually select polygons in arcgis or leaflet operation
2) intersect with the ahupuaʻa polygon and assign labels based on the ahupuaʻa that different plantations fall in


Running questions:
1) how do you set the CRS when the crs function continues to return null? does the crs() function always return null on an sf object? 

