---
title: "Patterns of Vegetation Recovery on Abandoned Agricultural Land"
author: "D. Nākoa Farrant"
date: '2022-06-29'
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

```{r}
library(plyr)
library(tidyverse) #for datafrmae manipulation
library(data.table) # for fread()
library(Hmisc)
library(corrplot)
library(dplyr)
library(sf)
library(lwgeom)
library(stargazer)
library(kableExtra)
library(viridis) # visualization package for color
library(RColorBrewer)
library(raster)
library(sp)
library(rgdal)
library(terra)
library(cowplot) # for ggdraw plot
library(patchwork) # new plotting package
library(ggspatial)
#library(tmap) # alternative to leaflet to do some interactive maps and maybe select polygons interactively 
#library(leaflet) # ideal for manual selection of polygons, but unsure if will have the opportunity to do that 
#library(maptools) # for zonal statistics
library(rnaturalearth)
library(rnaturalearthdata)
```


# Abandonment sequence output from google earth engine
```{r}
abS_envvar <- st_read("gee_files-abanSugar/abanS_mergeStats_rmvEarlyAbandon.shp") %>% 
  st_transform(st_crs(4326))%>% 
  st_set_precision(1e6) %>% 
  st_make_valid()
```


```{r}
abS_short <- abS_envvar %>% 
  dplyr::select(start, end, habstatus, lc, ndvi, evi, savi, msavi, ndmi, areaHa)
```

```{r}
HI_coastline <- st_read("coast_n83/coast_n83.shp") %>% 
  st_transform(st_crs(4326)) %>% 
  st_set_precision(1e6) %>% 
  st_make_valid() # class sf and data.frame
```

```{r}
ggplot(abS_clean) +
  geom_sf(aes(fill = start, color = start))

ggplot(abS_clean) +
  geom_sf(aes(fill = end, color = end))
```

```{r}
# Ran repair geometry on cah_land_cover_poly.shp in ArcMap on June 21, 2022
cah_lc_sf_repairedArc <- st_read("carbon_assessment_hi_shp/cah_repair_shp.shp") %>% # downloaded from GEE asset
  st_transform(st_crs(4326)) %>% 
  st_set_precision(1e6) %>% 
  st_make_valid() 
```

```{r}
cah_lc_sf_repairedArc <- cah_lc_sf_repairedArc %>% 
  tibble::rownames_to_column("FID") %>% 
  mutate(FID == as.numeric(FID))
```


```{r}
sf::sf_use_s2(FALSE)
abanS_rmvEarlyAban_cah <- st_intersection(st_make_valid(abS_envvar), st_make_valid(cah_lc_sf_repairedArc))

```

```{r}
abanS_rmvEarlyAban_cah <- abanS_rmvEarlyAban_cah %>% 
  mutate(FID = as.numeric(FID))
```

```{r}
abanS_rmvEarlyAban_cah_poly <- abanS_rmvEarlyAban_cah %>%
  filter(grepl("POLYGON", st_geometry_type(geometry)))
```


```{r}
#st_write(abanS_rmvEarlyAban_cah, "abandon_sugar_cah/abanS_rmvEarly_cah.gpkg", driver = "gpkg")
st_write(abanS_rmvEarlyAban_cah, "abandon_sugar_cah/abanS_rmvEarly_cah.shp")
# failed to create several features in writing the export
```



```{r}
abanS_cah_natural <- abanS_rmvEarlyAban_cah %>% 
  filter(Maj_LC == "Grassland" | Maj_LC == "Shrubland" | Maj_LC == "Forest" | Maj_LC == "Not Vegetated") %>% 
  tibble::rownames_to_column("ID") %>% 
  mutate(ID = as.numeric(ID))
```


```{r}
abanS_cah_natural_slim <- abanS_cah_natural %>% 
  dplyr::select(ID, startYr, endYr, areaHa, Gen_LC, Maj_LC, Det_LC, Hab_Status, Moisture_Z, Biome_Unit)
```

```{r}
abanS_cah_natural_slim_poly <- abanS_cah_natural_slim %>%
  filter(grepl("POLYGON", st_geometry_type(geometry)))
st_write(abanS_cah_natural_slim_poly, "abandon_sugar_cah/abanS_cah_natural_slim.shp")
```

# run this and the subsequent code chunks next
```{r}
hi_soil <- st_read("env_variables/soil_props/All_Properties.shp") %>% 
  st_transform(st_crs(4326)) %>% 
  st_set_precision(1e6) %>% 
  st_make_valid() # class sf and data.frame

# Geology spatial data source: https://www.sciencebase.gov/catalog/item/60df56d5d34ed15aa3b8a39c
hi_geology <- st_read("HI_geology/HawaiiStateGeologicMap_GeMS-open/GM_MapUnitPolys.shp") %>% 
  st_transform(st_crs(4326)) %>% 
  st_set_precision(1000000) %>% 
  st_make_valid() # class sf and data.frame

```


```{r}
# Add numeric columns for the min, max, and mean ages for different geological features
hi_geology <- hi_geology %>% 
  mutate(MinAge = as.numeric(ifelse(AgeRange == "1,500-3,000 yr", 1500, 
                          ifelse(AgeRange == "11,000-30,000 yr", 11000, 
                                 ifelse(AgeRange == "11,000-65,000 yr", 11000,
                                    ifelse(AgeRange == "120,000-260,000 yr", 120000,
                                           ifelse(AgeRange == "200-400 yr", 200,
                                                  ifelse(AgeRange == "200-750 yr", 200,
                                                         ifelse(AgeRange == "260,000-500,000 yr", 260000,
                                                                ifelse(AgeRange == "3,000-5,000 yr", 3000,
                                                                       ifelse(AgeRange == "30,000-50,000 yr", 30000,
                                                                              ifelse(AgeRange == "30,000-70,000 yr", 30000,
                                                                                     ifelse(AgeRange == "400-750 yr", 400,
                                                                                            ifelse(AgeRange == "5,000-11,000 yr", 5000,
                                                                                                   ifelse(AgeRange == "64,000-300,000 yr", 64000,
                                                                                                          ifelse(AgeRange == "750-1,500 yr", 750,
                                                                                                                 ifelse(AgeRange == "A.D. 1410-1460", 2020-1460,
                                                                                                                        ifelse(AgeRange == "A.D. 1790", 2020-1790,
                                                                                                                               ifelse(AgeRange == "A.D. 1823", 2020-1823,
                                                                                                                                      ifelse(AgeRange == "A.D. 1840", 2020-1840,
                                                                                                                                             ifelse(AgeRange == "A.D. 1868", 2020-1868,
                                                                                                                                                    ifelse(AgeRange == "A.D. 1880-1881", 2020-1881,
                                                                                                                                                           ifelse(AgeRange == "A.D. 1955", 2020-1955,
                                                                                                                                                                  ifelse(AgeRange == "A.D. 1961", 2020-1961,
                                                                                                                                                                                ifelse(AgeRange == "A.D. 2007-2018", 2020-2018,
                                                                                                                                                                                       ifelse(AgeRange == "A.D. 2018", 2020-2018,
                                      ifelse(AgeRange == "Age 0.1-0.2 Ma", 100000, "NA"))))))))))))))))))))))))))) %>% 
  mutate(MaxAge = as.numeric(ifelse(AgeRange == "1,500-3,000 yr", 3000, 
                          ifelse(AgeRange == "11,000-30,000 yr", 30000, 
                                 ifelse(AgeRange == "11,000-65,000 yr", 65000,
                                    ifelse(AgeRange == "120,000-260,000 yr", 260000,
                                           ifelse(AgeRange == "200-400 yr", 400,
                                                  ifelse(AgeRange == "200-750 yr", 750,
                                                         ifelse(AgeRange == "260,000-500,000 yr", 500000,
                                                                ifelse(AgeRange == "3,000-5,000 yr", 5000,
                                                                       ifelse(AgeRange == "30,000-50,000 yr", 50000,
                                                                              ifelse(AgeRange == "30,000-70,000 yr", 70000,
                                                                                     ifelse(AgeRange == "400-750 yr", 750,
                                                                                            ifelse(AgeRange == "5,000-11,000 yr", 11000,
                                                                                                   ifelse(AgeRange == "64,000-300,000 yr", 300000,
                                                                                                          ifelse(AgeRange == "750-1,500 yr", 1500,
                                                                                                                 ifelse(AgeRange == "A.D. 1410-1460", 2020-1410,
                                                                                                                        ifelse(AgeRange == "A.D. 1790", 2020-1790,
                                                                                                                               ifelse(AgeRange == "A.D. 1823", 2020-1823,
                                                                                                                                      ifelse(AgeRange == "A.D. 1840", 2020-1840,
                                                                                                                                             ifelse(AgeRange == "A.D. 1868", 2020-1868,
                                                                                                                                                    ifelse(AgeRange == "A.D. 1880-1881", 2020-1880,
                                                                                                                                                           ifelse(AgeRange == "A.D. 1955", 2020-1955,
                                                                                                                                                                  ifelse(AgeRange == "A.D. 1961", 2020-1961,
                                                                                                                                                                                ifelse(AgeRange == "A.D. 2007-2018", 2020-2007,
                                                                                                                                                                                       ifelse(AgeRange == "A.D. 2018", 2020-2018,
                                      ifelse(AgeRange == "Age 0.1-0.2 Ma", 200000, "NA"))))))))))))))))))))))))))) %>% 
    mutate(MeanAge = as.numeric(ifelse(AgeRange == "1,500-3,000 yr", 2250, 
                          ifelse(AgeRange == "11,000-30,000 yr", 20500, 
                                 ifelse(AgeRange == "11,000-65,000 yr", 38000,
                                    ifelse(AgeRange == "120,000-260,000 yr", 190000,
                                           ifelse(AgeRange == "200-400 yr", 300,
                                                  ifelse(AgeRange == "200-750 yr", 475,
                                                         ifelse(AgeRange == "260,000-500,000 yr", 380000,
                                                                ifelse(AgeRange == "3,000-5,000 yr", 4000,
                                                                       ifelse(AgeRange == "30,000-50,000 yr", 40000,
                                                                              ifelse(AgeRange == "30,000-70,000 yr", 50000,
                                                                                     ifelse(AgeRange == "400-750 yr", 575,
                                                                                            ifelse(AgeRange == "5,000-11,000 yr", 8000,
                                                                                                   ifelse(AgeRange == "64,000-300,000 yr", 182000,
                                                                                                          ifelse(AgeRange == "750-1,500 yr", 1125,
                                                                                                                 ifelse(AgeRange == "A.D. 1410-1460", 2020-1435,
                                                                                                                        ifelse(AgeRange == "A.D. 1790", 2020-1790,
                                                                                                                               ifelse(AgeRange == "A.D. 1823", 2020-1823,
                                                                                                                                      ifelse(AgeRange == "A.D. 1840", 2020-1840,
                                                                                                                                             ifelse(AgeRange == "A.D. 1868", 2020-1868,
                                                                                                                                                    ifelse(AgeRange == "A.D. 1880-1881", 2020-1880,
                                                                                                                                                           ifelse(AgeRange == "A.D. 1955", 2020-1955,
                                                                                                                                                                  ifelse(AgeRange == "A.D. 1961", 2020-1961,
                                                                                                                                                                                ifelse(AgeRange == "A.D. 2007-2018", 2020-2012,
                                                                                                                                                                                       ifelse(AgeRange == "A.D. 2018", 2020-2018,
                                      ifelse(AgeRange == "Age 0.1-0.2 Ma", 150000, "NA")))))))))))))))))))))))))))
  
```

```{r}
# intersect with soil properties
abanS_rmvEarlyAban_cah_soil <- st_intersection(st_make_valid(abanS_cah_natural_slim_poly), st_make_valid(hi_soil))
# intersect with geology properties
abanS_rmvEarlyAban_cah_geology <- st_intersection(st_make_valid(abanS_cah_natural_slim_poly), st_make_valid(hi_geology))

abanS_rmvEarlyAban_cah_soil_geology <- st_intersection(st_make_valid(abanS_rmvEarlyAban_cah_soil), st_make_valid(hi_geology))
```


```{r}


abanS_cah_natural_slim_poly <- abanS_cah_natural_slim %>%
  filter(grepl("POLYGON", st_geometry_type(geometry)))
st_write(abanS_cah_natural_slim_poly, "abandon_sugar_cah/abanS_cah_natural_slim.shp")
```

# General abandoned sugarcane plots
```{r}
###world
world <- ne_countries(scale = "medium", returnclass = "sf") %>% 
  st_transform(st_crs(4326))
```

```{r}
main_map <- ggplot() +
  geom_sf(data = HI_coastline) + 
  geom_sf(data = abS_clean, aes(fill = start, color = start)) + 
  coord_sf(xlim = c(-160.5, -154.2), ylim = c(18.5, 22.8), expand = FALSE) +
  scale_fill_brewer(palette="YlOrRd") +
  scale_color_brewer(palette="YlOrRd") +
  labs(x = "Longitude (º)", y = "Latitude (º)") +
  theme_bw()+
  theme(panel.border = element_rect(colour = "black", fill=NA, size=1),
  panel.background = element_rect(fill = "#2A788EFF"), panel.grid.major = element_line(size = 0.5, linetype = 'solid',
                                colour = "#2A788EFF"))+
  annotation_scale(location = "br")

main_map
```

```{r}
main_map + 
  coord_sf(xlim = c(-159.9, -159.2), ylim = c(21.8, 22.3), expand = FALSE)
```

```{r}
# not a great figure
ggdraw(main_map) +
  draw_plot(
    {
      main_map + 
      coord_sf(xlim = c(-159.9, -159.2), ylim = c(21.8, 22.3), expand = FALSE) +
      theme(legend.position = "none")
    },
    x = 0, 
    # The distance along a (0,1) y-axis to draw the bottom edge of the plot
    y = 0.6,
    # The width and height of the plot expressed as proportion of the entire ggdraw object
    width = 0.4, 
    height = 0.4
  )
```

Kauai plot
```{r}
ggplot() +
  geom_sf(data = HI_coastline) + 
  geom_sf(data = abS_clean, aes(fill = start, color = start)) + 
  coord_sf(xlim = c(-159.9, -159.2), ylim = c(21.8, 22.3), expand = FALSE) +
  #coord_sf(xlim = c(-160.5, -154.2), ylim = c(18.5, 22.8), expand = FALSE) +
  #geom_point(data = sites, aes(x = longitude, y = latitude), size = 4, 
  #           shape = 23, fill = "black",size=2) +
  # annotate("text", x = -122.103777, y = 51.179863, label = "Big Bar Landslide", angle = 45) +
  #geom_label(data=sites, aes(x=longitude, y=latitude, label=name), 
  #           nudge_x=0.17, nudge_y=0.15, label.size=0.1,size=3, 
  #           fontface = "bold.italic", label.r=unit(0.20, "lines"))+
  scale_fill_brewer(palette="YlOrRd") +
  scale_color_brewer(palette="YlOrRd") +
  labs(x = "Longitude (º)", y = "Latitude (º)") +
  theme_bw()+
  theme(panel.border = element_rect(colour = "black", fill=NA, size=1),
  panel.background = element_rect(fill = "#2A788EFF"), panel.grid.major = element_line(size = 0.5, linetype = 'solid',
                                colour = "#2A788EFF"))+
  annotation_scale(location = "br")
```

Oahu
```{r}
ggplot() +
  geom_sf(data = HI_coastline) + 
  geom_sf(data = abS_clean, aes(fill = start, color = start)) + 
  coord_sf(xlim = c(-158.5, -157.5), ylim = c(21.1, 21.9), expand = FALSE) +
  labs(x = "Longitude (º)", y = "Latitude (º)") +
  scale_fill_brewer(palette="YlOrRd") +
  scale_color_brewer(palette="YlOrRd") +
  theme_bw()+
  theme(panel.border = element_rect(colour = "black", fill=NA, size=1),
  panel.background = element_rect(fill = "#2A788EFF"), panel.grid.major = element_line(size = 0.5, linetype = 'solid',
                                colour = "#2A788EFF"))+
  annotation_scale(location = "br")
```

```{r}
sf::sf_use_s2(FALSE)
```


```{r}
abS_clean$geometry <- abS_clean$geometry %>%
  s2::s2_rebuild() %>%
  sf::st_as_sfc()
```



```{r}
# necessary to avoid error with the st_crop calls s2_geography_from_wkb(x, oriented = oriented, check = check)
sf::sf_use_s2(FALSE)

kauai_box <- c(xmin= -160, ymin = 21.6, xmax = -159, ymax = 22.4) # 1 x .8

oahu_box <- c(xmin= -158.5, ymin = 20.9, xmax = -157.5, ymax = 21.9) # 1 x 1

maui_box <- c(xmin= -156.7, ymin = 20.3, xmax = -155.9, ymax = 21.05) # 0.8 x 0.7

hawaii_box <- c(xmin= -156.5, ymin = 18.8, xmax = -154.0, ymax = 20.5) # 1.7 x 1.7

kauai_coast <- st_crop(HI_coastline, kauai_box)
oahu_coast <- st_crop(HI_coastline, oahu_box)
maui_coast <- st_crop(HI_coastline, maui_box)
hawaii_coast <- st_crop(HI_coastline, hawaii_box)

kauai_crop <- st_crop(abS_clean, kauai_box)
oahu_crop <- st_crop(abS_clean, oahu_box)
maui_crop <- st_crop(abS_clean, maui_box)
hawaii_crop <- st_crop(st_make_valid(abS_clean), hawaii_box)

```

# Add an Island column
```{r}
kauai <- kauai_crop %>% 
  mutate(Island = "Kauai")
oahu <- oahu_crop %>% 
  mutate(Island = "Oahu")
maui <- maui_crop %>% 
  mutate(Island = "Maui")
hawaii <- hawaii_crop %>% 
  mutate(Island = "Hawaii") 

kauai_coast <- kauai_coast %>% 
  mutate(Island = "Kauai")
oahu_coast <- oahu_coast %>% 
  mutate(Island = "Oahu")
maui_coast <- maui_coast %>% 
  mutate(Island = "Maui")
hawaii_coast <- hawaii_coast %>% 
  mutate(Island = "Hawaii") 

```

# Join the different Island shapefiles that all have the same columns so can bind the rows
```{r}
ka_oa <- rbind(kauai, oahu)
ka_oa_ma <- rbind(ka_oa, maui)
ka_oa_ma_ha <- rbind(ka_oa_ma, hawaii)

ka_oa_coast <- rbind(kauai_coast, oahu_coast)
ka_oa_ma_coast <- rbind(ka_oa_coast, maui_coast)
ka_oa_ma_ha_coast <- rbind(ka_oa_ma_coast, hawaii_coast)
```



```{r}
ggplot() +
  geom_sf(data = ka_oa_ma_ha_coast) + 
  geom_sf(data = ka_oa_ma_ha, aes(fill = start, color = start)) +
  facet_wrap(~Island) +
  #coord_equal() +
  labs(x = "Longitude (º)", y = "Latitude (º)") +
  scale_fill_brewer(palette="YlOrRd") +
  scale_color_brewer(palette="YlOrRd") +
  theme_bw()+
  theme(panel.border = element_rect(colour = "black", fill=NA, size=1),
  panel.background = element_rect(fill = "#2A788EFF"), panel.grid.major = element_line(size = 0.5, linetype = 'solid',
                                colour = "#2A788EFF")) +
  annotation_scale(location = "br")
```


# Plots using the Patchwork package
```{r}

kauai_plot <- ggplot() +
  geom_sf(data = kauai_coast) + 
  geom_sf(data = kauai_crop, aes(fill = start, color = start)) + 
  coord_sf(xlim = c(-159.9, -159.1), ylim = c(21.75, 22.4), expand = FALSE) +
  labs(x = "Longitude (º)", y = "Latitude (º)") +
  scale_fill_brewer(palette="YlOrRd") +
  scale_color_brewer(palette="YlOrRd") +
  theme_bw()+
  theme(legend.position = "none", panel.border = element_rect(colour = "black", fill=NA, size=1),
  panel.background = element_rect(fill = "#2A788EFF"), panel.grid.major = element_line(size = 0.5, linetype = 'solid',
                                colour = "#2A788EFF"))+
  annotation_scale(location = "br")

oahu_plot <- ggplot() +
  geom_sf(data = oahu_coast) + 
  geom_sf(data = oahu_crop, aes(fill = start, color = start)) +
  coord_sf(xlim = c(-158.4, -157.5), ylim = c(21.15, 21.8), expand = FALSE) +
  labs(x = "Longitude (º)", y = "Latitude (º)") +
  scale_fill_brewer(palette="YlOrRd") +
  scale_color_brewer(palette="YlOrRd") +
  theme_bw()+
  theme(legend.position = "none", panel.border = element_rect(colour = "black", fill=NA, size=1),
  panel.background = element_rect(fill = "#2A788EFF"), panel.grid.major = element_line(size = 0.5, linetype = 'solid',
                                colour = "#2A788EFF"))+
  annotation_scale(location = "br")

maui_plot <- ggplot() +
  geom_sf(data = maui_coast) + 
  geom_sf(data = maui_crop, aes(fill = start, color = start)) + 
  coord_sf(xlim = c(-156.7, -155.9), ylim = c(20.5, 21.05), expand = FALSE) +
  labs(x = "Longitude (º)", y = "Latitude (º)") +
  scale_fill_brewer(palette="YlOrRd") +
  scale_color_brewer(palette="YlOrRd") +
  theme_bw()+
  theme(legend.position = "none", panel.border = element_rect(colour = "black", fill=NA, size=1),
  panel.background = element_rect(fill = "#2A788EFF"), panel.grid.major = element_line(size = 0.5, linetype = 'solid',
                                colour = "#2A788EFF"))+
  annotation_scale(location = "br")

hawaii_plot <- ggplot() +
  geom_sf(data = hawaii_coast) + 
  geom_sf(data = hawaii_crop, aes(fill = start, color = start)) + 
  labs(x = "Longitude (º)", y = "Latitude (º)") +
  coord_sf(xlim = c(-156.4, -154.4), ylim = c(18.8, 20.4), expand = FALSE) +
  scale_fill_brewer(palette="YlOrRd") +
  scale_color_brewer(palette="YlOrRd") +
  theme_bw()+
  theme(legend.position = "none", panel.border = element_rect(colour = "black", fill=NA, size=1),
  panel.background = element_rect(fill = "#2A788EFF"), panel.grid.major = element_line(size = 0.5, linetype = 'solid',
                                colour = "#2A788EFF"))+
  annotation_scale(location = "br")

```

```{r}
# also need to look into cowplot in addition to patchwork
# https://jayrobwilliams.com/posts/2021/05/geom-sf-facet
kauai_plot + oahu_plot + maui_plot + hawaii_plot
```


```{r}
hawaii_long <- gather(hawaii, condition, measurement, habstatus:areaHa) 
  
```

```{r}
hawaii_long_interest <- hawaii_long %>% 
  filter(condition == "ndvi" | condition == "evi" | condition == "msavi" | condition == "ndmi")
```

```{r}
ggplot() +
  geom_sf(data = hawaii_long_interest, aes(color = measurement, fill = measurement)) +
  scale_y_continuous(breaks = seq(18.8, 20.4, by = 0.2)) +
  scale_x_continuous(breaks = seq(-156.4, -154.4, by = 0.2)) +
  facet_wrap(~hawaii_long_interest$condition)
```



```{r}
abS_cah_natural_stats <- st_read("gee_files-abanSugar/abS_cah_natural_stats.shp")
```



# Plots
```{r}
reserves <- st_read("env_reserves_hi/reserves.shp") %>% 
  st_transform(st_crs(4326)) %>% 
  st_set_precision(1000000) %>% 
  st_make_valid() # class sf and data.frame
```

```{r}
abS_cah_natural_stats_reserves <- st_intersection(abS_cah_natural_stats, reserves)
```

```{r}
summary(abS_cah_natural_stats_reserves$NAME %>% as.factor())
```



```{r}
abS_cah_natural_stats_reserves <- abS_cah_natural_stats_reserves %>% 
  mutate(SurveyYrAb = endYr)

abS_cah_natural_stats_reserves$SurveyYrAb[abS_cah_natural_stats_reserves$SurveyYrAb == 2020] <- 2016
abS_cah_natural_stats_reserves$SurveyYrAb[abS_cah_natural_stats_reserves$SurveyYrAb == 2015] <- 1995

abS_cah_natural_stats_reserves$TimeSinceAb <- 2020-abS_cah_natural_stats_reserves$SurveyYrAb

abS_cah_natural_stats_reserves <- abS_cah_natural_stats_reserves %>% 
  mutate(MaxCultTime = ifelse(startYr == 1900, SurveyYrAb - 1900, 
                          ifelse(startYr == 1920, SurveyYrAb - 1901, 
                                 ifelse(startYr == 1937, SurveyYrAb - 1921,
                                    ifelse(startYr == 1980, SurveyYrAb - 1938,
                                      ifelse(startYr == 2015, SurveyYrAb - 1981, SurveyYrAb - 2016))))))
```


# example plots for plotting variation in vegetation indices
```{r}
# example plots for plotting variation in vegetation indices
ggplot(abS_cah_natural_stats_reserves, aes(x = TimeSinceAb, y=evi)) + 
  geom_point(aes(color = Hab_Status))+
  #geom_point(color = "orange") +
  facet_wrap(~ Hab_Status) +
  theme_bw() +
  xlab("") +
  ylab("Variance in EVI") +
  labs(color = "") +
  theme(text = element_text(size=18),
    legend.justification = c("right", "bottom"),
    legend.background = element_rect(colour = "black", size = 0.1),
           legend.key.size = unit(1, "lines")) #+
  #scale_color_manual(values = c("purple", "darkblue", "cyan", "deeppink1"))

ggplot(abS_cah_natural_stats_reserves %>% filter(Maj_LC != "Not Vegetated"), aes(x = TimeSinceAb, y=ndmi)) + 
  geom_point(aes(color = Hab_Status))+
  #geom_point(color = "orange") +
  facet_wrap(~ Hab_Status)
  theme_bw() +
  xlab("") +
  ylab("Variance in NDMI") +
  labs(color = "") #+
  theme(text = element_text(size=18),
    legend.justification = c("right", "bottom"),
    legend.background = element_rect(colour = "black", size = 0.1),
           legend.key.size = unit(1, "lines")) #+
  #scale_color_manual(values = c("purple", "darkblue", "cyan", "deeppink1"))
  
ggplot(abS_cah_natural_stats_reserves %>% filter(Maj_LC != "Not Vegetated", abs(evi) < 2), aes(x = TimeSinceAb, y=ndmi)) + 
  geom_point(aes(color = Hab_Status))+
  #geom_point(color = "orange") +
  facet_wrap(~ Maj_LC) +
  theme_bw() +
  xlab("") +
  ylab("NDRE") +
  labs(color = "") #+
  theme(text = element_text(size=18),
    legend.justification = c("right", "bottom"),
    legend.background = element_rect(colour = "black", size = 0.1),
           legend.key.size = unit(1, "lines")) #+
  #scale_color_manual(values = c("purple", "darkblue", "cyan", "deeppink1"))
```

# biomass sample plots 
```{r}
ggplot(abS_cah_natural_stats_reserves %>% filter(Maj_LC != "Not Vegetated"), aes(x = TimeSinceAb, y=MEAN)) + 
  geom_point(aes(color = Year)) +  
  #geom_point(color = "orange") + 
  geom_smooth(method = "lm", color = "orange") +
  geom_smooth(method = "lm", formula = y ~ log(x), color = "darkgreen") +
  facet_wrap(~ Moisture_Z) +
  theme_bw() +
  xlab("Years since abandonment") +
  ylab(bquote('Biomass (Mg C/ha)')) +
  labs(color = "Cultivation Began") +
  theme(text = element_text(size=18),
    legend.justification = c("right", "bottom"),
    legend.background = element_rect(colour = "black", size = 0.1),
           legend.key.size = unit(1, "lines")) #+
  #scale_color_manual(values = c("purple", "darkblue", "cyan", "deeppink1"))

```


```{r}

summarizedAreaByTimeSinceAb <- abS_cah_natural_stats_reserves %>% 
  #filter(Maj_LC != "Not Vegetated") %>% 
  group_by(TimeSinceAb) %>% 
  summarise(AbBinArea = sum(areaHa)) %>% 
  arrange(desc(TimeSinceAb)) %>%
  mutate(prop = AbBinArea / sum(AbBinArea) *100)

summarizeAreaByStartYear <- abS_cah_natural_stats_reserves %>%
  #filter(Hab_Status != "Bare < 5% plant cover") %>% 
  group_by(startYr) %>% 
  summarise(AbBinArea = sum(areaHa)) %>% 
  ungroup() %>% 
  arrange(desc(startYr)) %>%
  mutate(prop = AbBinArea / sum(AbBinArea) *100)
```



```{r}
ggplot(summarizedAreaByTimeSinceAb, aes(x = "", y= AbBinArea, fill = as.factor(TimeSinceAb))) +
  geom_bar(stat="identity", width=1) +
  coord_polar("y", start=0) +
  theme_void() +
  labs(fill = "Time Since Abandonment") +
  theme(text = element_text(size=18),
    legend.justification = c("right", "bottom"),
    legend.background = element_rect(colour = "black", size = 0.1),
           legend.key.size = unit(1, "lines")) +
  scale_fill_manual(values = c("purple", "darkblue", "cyan", "deeppink1", "red1", "orange"))

ggplot(summarizeAreaByStartYear, aes(x = "", y= AbBinArea, fill = as.factor(startYr))) +
  geom_bar(stat="identity", width=1) +
  coord_polar("y", start=0) +
  theme_void() +
  labs(fill = "Initial Year Cultivated") +
  theme(text = element_text(size=18),
    legend.justification = c("right", "bottom"),
    legend.background = element_rect(colour = "black", size = 0.1),
           legend.key.size = unit(1, "lines")) 
```


# Plots for vegetation cover (Maj_LC)
```{r}
abS_cah_natural_stats_reserves$Maj_LC <- factor(abS_cah_natural_stats_reserves$Maj_LC, levels = c("Forest", "Shrubland", "Grassland", "Not Vegetated"))
```

```{r}
ggplot(abS_cah_natural_stats_reserves, aes(fill=Maj_LC, y=areaHa, x=TimeSinceAb %>% as.factor())) + 
    geom_bar(position="stack", stat="identity") + 
  theme_bw() +
  xlab("Years since abandonment") +
  ylab(bquote('Total Area (ha)')) +
  labs(fill = "Land Cover") +
  theme(text = element_text(size=18),
    legend.justification = c("right", "bottom"),
    legend.background = element_rect(colour = "black", size = 0.1),
           legend.key.size = unit(1, "lines")) + scale_fill_manual(values = c("darkgreen", "green2", "lightgreen", "sienna"))

ggplot(abS_cah_natural_stats_reserves, aes(fill=Maj_LC, y=areaHa, x=TimeSinceAb %>% as.factor())) + 
    geom_bar(position="fill", stat="identity") + 
  theme_bw() +
  xlab("Years since abandonment") +
  ylab("Proportion of Area") +
  labs(fill = "Land Cover") +
  theme(text = element_text(size=18),
    legend.justification = c("right", "bottom"),
    legend.background = element_rect(colour = "black", size = 0.1),
           legend.key.size = unit(1, "lines")) + scale_fill_manual(values = c("darkgreen", "green2", "lightgreen", "sienna"))

ggplot(abS_cah_natural_stats_reserves %>% filter(Maj_LC != "Not Vegetated"), aes(fill=Maj_LC, y=areaHa, x=TimeSinceAb %>% as.factor())) + 
    geom_bar(position="stack", stat="identity") + 
  theme_bw() +
  xlab("Years since abandonment") +
  ylab(bquote('Total Area (ha)')) +
  labs(fill = "Land Cover") +
  theme(text = element_text(size=18),
    legend.justification = c("right", "bottom"),
    legend.background = element_rect(colour = "black", size = 0.1),
           legend.key.size = unit(1, "lines")) + scale_fill_manual(values = c("darkgreen", "green2", "lightgreen", "sienna"))

ggplot(abS_cah_natural_stats_reserves %>% filter(Maj_LC != "Not Vegetated"), aes(fill=Maj_LC, y=areaHa, x=TimeSinceAb %>% as.factor())) + 
    geom_bar(position="fill", stat="identity") + 
  theme_bw() +
  xlab("Years since abandonment") +
  ylab("Proportion of Total Area") +
  labs(fill = "Land Cover") +
  theme(text = element_text(size=18),
    legend.justification = c("right", "bottom"),
    legend.background = element_rect(colour = "black", size = 0.1),
           legend.key.size = unit(1, "lines")) + scale_fill_manual(values = c("darkgreen", "green2", "lightgreen", "sienna"))
```


# Plots for vegetation composition (HabStatus)
```{r}
abS_cah_natural_stats_reserves$Hab_Status <- factor(abS_cah_natural_stats_reserves$Hab_Status, levels = c("Native Dominated", "Native / Alien Mix", "Bare < 5% plant cover"))
```

```{r}
ggplot(abS_cah_natural_stats_reserves, aes(fill=Hab_Status, y=areaHa, x=TimeSinceAb %>% as.factor())) + 
    geom_bar(position="stack", stat="identity") + 
  theme_bw() +
  xlab("Years since abandonment") +
  ylab(bquote('Total Area (ha)')) +
  labs(fill = "Land Cover") +
  theme(text = element_text(size=18),
    legend.justification = c("right", "bottom"),
    legend.background = element_rect(colour = "black", size = 0.1),
           legend.key.size = unit(1, "lines")) + scale_fill_manual(values = c("darkgreen", "orange", "sienna"), labels = c("Native", "Exotic", "Not Vegetated"))

ggplot(abS_cah_natural_stats_reserves, aes(fill=Hab_Status, y=areaHa, x=TimeSinceAb %>% as.factor())) + 
    geom_bar(position="fill", stat="identity") + 
  theme_bw() +
  xlab("Years since abandonment") +
  ylab("Proportion of Area") +
  labs(fill = "Land Cover") +
  theme(text = element_text(size=18),
    legend.justification = c("right", "bottom"),
    legend.background = element_rect(colour = "black", size = 0.1),
           legend.key.size = unit(1, "lines")) + scale_fill_manual(values = c("darkgreen", "orange", "sienna"), labels = c("Native", "Exotic", "Not Vegetated"))

ggplot(abS_cah_natural_stats_reserves %>% filter(Hab_Status != "Bare < 5% plant cover"), aes(fill=Hab_Status, y=areaHa, x=TimeSinceAb %>% as.factor())) + 
    geom_bar(position="stack", stat="identity") + 
  #facet_wrap(~ startYr) +
  theme_bw() +
  xlab("Years since abandonment") +
  ylab(bquote('Total Area ('*'km'^2*')')) +
  labs(fill = "Land Cover") +
  theme(text = element_text(size=18),
    legend.justification = c("right", "bottom"),
    legend.background = element_rect(colour = "black", size = 0.1),
           legend.key.size = unit(1, "lines")) + scale_fill_manual(values = c("darkgreen", "orange"), labels = c("Native", "Exotic"))
```

```{r}
MixHabitatStatus <- abS_cah_natural_stats_reserves %>% filter(Hab_Status == "Native / Alien Mix")

summary(MixHabitatStatus$Det_LC) # basically all of the Mixed habitats are labeled as alien species in the Detailed LC other than a few pixels that are labeled as mixed native-alien mesic forest

```


```{r}
abS_cah_natural_stats_reserves$VegCover = ifelse(abS_cah_natural_stats_reserves$Hab_Status == "Native Dominated", "Native", 
                                                         ifelse(abS_cah_natural_stats_reserves$Hab_Status == "Native / Alien Mix", "Invasive", "Bare"))
```



# To Do's
- ideally with the vegetation index extractions, will want the variance of those different vegetation indices within the different areas. Do you want that variance even if you aren't doing seasonal assessments
- retain more of the CAH properties 
- retain the certain soil properties in the shape you upload?
- intesect with soil shapefile before doing the extractions on GEE?
- include precipitation, intersect with soil shape? before doing the extraction? 
- create the counterfactual shapes. Intersect with CAH. extract the same vegetation indices, elevation, slope, aspect, precipitation, 