---
title: "Sugar Species Distribution Model"
author: "D. Nākoa Farrant"
date: '2022-06-29'
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

```{r}
library(tidyverse)
library(RColorBrewer)
library(cowplot) # for ggdraw plot
library(raster)
library(fasterize)
library(sf)
library(rgdal)
library(terra)
library(wallace)
```

Goal is to run a species distribution model (Maxent) to determine the habitat suitability envelope for sugarcane across Hawai'i
1) Get "occurrence" points by gridding the sugarcane shapefiles for all historic sugarcane in Hawai'i. Make sure it's in the same projection as all of the other environmental data
2) Gather environmental data:
- precipitation, temperature, elevation, slope, aspect?, soil organic matter percent, soil cation exchange capacity

3) ensure all environmental rasters are in the same projection, extent, and resolution

4) specify a region of interest

5) lay down random background points?

```{r}
hi_soil <- st_read("env_variables/soil_props/All_Properties.shp") %>% 
  st_transform(st_crs(4326)) %>% 
  st_set_precision(1e6) %>% 
  st_make_valid() # class sf and data.frame

```


```{r}
ext <- floor(extent(hi_soil))
r <- raster(ext, res = 0.00225) # resolution taken from the precipiation raster
cec_raster <- fasterize(hi_soil, r, "CECmod_rep")# fasterize is way faster than rasterize function
pH_raster <- fasterize(hi_soil, r, "pHmod_rep")
Ksat_raster <- fasterize(hi_soil, r, "Ksat_rep")
OM_raster <- fasterize(hi_soil, r, "OM_rep")
```

```{r}
#plot(cec_raster)
#plot(pH_raster)
#plot(Ksat_raster)
#plot(OM_raster)
```

```{r}
# Landfire Topographical data for 2020
# https://landfire.gov/version_download.php#
# Elevation units of meters from 0m to 4199m
# grid cell is 30 meters 
elev <- raster("env_variables/landfire_topography/LH20_elev_GCS_WGS84.tif")

# Aspect units of degrees from 0 to 359
# grid cell is 30 meters
# Per Carla's previous recommendation, might need to convert from degrees to be some kind of heading like north reading
aspect <- raster("env_variables/landfire_topography/LH20_aspect_GCS_WGS84.tif")

# Slope degrees from 0 to 79 degree
# grid cells is 30 meters
slope_deg <- raster("env_variables/landfire_topography/LH20_slpdeg_GCS_WGS84.tif")

# Mean annual precipitation 250m resolution from HI rainfall atlas
# mm units
# http://rainfall.geography.hawaii.edu/downloads.html
precip <- raster("env_variables/StateMapTIFs_mm/staterf_mmann_GCS_WGS84.tif") 

# Mean annual air temperature 250m resolution from HI Evapotranspiration atlas
# degree C units
temp <- raster("env_variables/Tair_month_raster/tair_ann.tif")

# Relative humidity annually at 250m resolution 
# http://evapotranspiration.geography.hawaii.edu/downloads.html
humidity <- raster("env_variables/RH_month_raster/rh_ann.tif")

# soil moisture at 250m resolution
# http://evapotranspiration.geography.hawaii.edu/downloads.html
soil_moist <- raster("env_variables/SoilMoisture_month_raster/sl_mst_ann.tif")

# Surface temperature annually at 250m. Degrees C
# http://evapotranspiration.geography.hawaii.edu/downloads.html
surf_temp <- raster("env_variables/TSurf_month_raster/tsurf_ann.tif")

# read in slope percent rise raster
slp_pct <- raster("env_variables/landfire_topography/LH20_slppct_GCS_WGS84.tif")
```

```{r}
biomass_wgs84 <- raster("env_variables/Biomass/asner_biomass_30m_WGS84.tif")
```


```{r}
aspect_radians <- aspect*pi/180 # convert from degrees to radians

asp_northing <- cos(aspect_radians) # extract aspect northness component
asp_easting <- sin(aspect_radians) # extract aspect eastness component
```


# Project all of the environmentl rasters into the same projection
```{r}
crs(elev) <- st_crs(4326)$wkt
crs(aspect) <- st_crs(4326)$wkt
crs(slope_deg) <- st_crs(4326)$wkt
crs(precip) <- st_crs(4326)$wkt
crs(temp) <- st_crs(4326)$wkt
crs(humidity) <- st_crs(4326)$wkt
crs(soil_moist) <- st_crs(4326)$wkt
crs(surf_temp) <- st_crs(4326)$wkt
crs(cec_raster) <- st_crs(4326)$wkt
crs(pH_raster) <- st_crs(4326)$wkt
crs(Ksat_raster) <- st_crs(4326)$wkt
crs(OM_raster) <- st_crs(4326)$wkt
```


# check resolutin of environmental layers
```{r}
res(elev)
res(aspect)
res(slope_deg)
res(slp_pct)
res(precip)
res(temp)
res(humidity)
res(soil_moist)
res(surf_temp)
res(cec_raster)
res(pH_raster)
res(Ksat_raster)
res(OM_raster)
```

```{r}

# method for resampling is either bilinear or nearest neighbor "ngb"
# nearest neighbor is best for categorical data
# description of resampling methods is here: https://support.esri.com/en/technical-article/000005606

elev_resamp <- resample(elev, precip, method = "bilinear")
slpdg_resamp <- resample(slope_deg, precip, method = "bilinear")
northing_resamp <- resample(asp_northing, precip, method = "bilinear")
easting_resamp <- resample(asp_easting, precip, method = "bilinear")
slppct_resamp <- resample(slp_pct, precip, method = "bilinear")
cec_resamp <- resample(cec_raster, precip, method = "bilinear")
pH_resamp <- resample(pH_raster, precip, method = "bilinear")
Ksat_resamp <- resample(Ksat_raster, precip, method = "bilinear")
OM_resamp <- resample(OM_raster, precip, method = "bilinear")
```

```{r}
biomass_resamp <- resample(biomass_wgs84, precip, method = "bilinear")
```


```{r}
extent(elev_resamp)
extent(slpdg_resamp)
extent(slppct_resamp)
extent(precip)
extent(temp)
extent(humidity)
extent(soil_moist)
extent(surf_temp)
extent(northing_resamp)
extent(easting_resamp)
extent(cec_resamp)
extent(pH_resamp)
extent(Ksat_resamp)
extent(OM_resamp)
```


```{r}
writeRaster(elev_resamp,"env_variables/sdm_env_var/elev.tif", format="GTiff", overwrite = T)
writeRaster(slpdg_resamp,"env_variables/sdm_env_var/slpdg.tif", format="GTiff", overwrite = T)
writeRaster(precip,"env_variables/sdm_env_var/precip.tif", format="GTiff", overwrite = T)
writeRaster(temp,"env_variables/sdm_env_var/temp.tif", format="GTiff", overwrite = T)
writeRaster(humidity,"env_variables/sdm_env_var/humid.tif", format="GTiff", overwrite = T)
writeRaster(soil_moist,"env_variables/sdm_env_var/soil_moist.tif", format="GTiff", overwrite = T)
writeRaster(surf_temp,"env_variables/sdm_env_var/surf_temp.tif", format="GTiff", overwrite = T)
writeRaster(northing_resamp,"env_variables/sdm_env_var/north.tif", format="GTiff", overwrite = T)
writeRaster(easting_resamp,"env_variables/sdm_env_var/east.tif", format="GTiff", overwrite = T)
writeRaster(slppct_resamp,"env_variables/sdm_env_var/slppct.tif", format="GTiff", overwrite = T)
writeRaster(cec_resamp,"env_variables/sdm_env_var/cec.tif", format="GTiff", overwrite = T)
writeRaster(pH_resamp,"env_variables/sdm_env_var/pH.tif", format="GTiff", overwrite = T)
writeRaster(Ksat_resamp,"env_variables/sdm_env_var/Ksat.tif", format="GTiff", overwrite = T)
writeRaster(OM_resamp,"env_variables/sdm_env_var/OM.tif", format="GTiff", overwrite = T)

```

```{r}
# Might not use biomass as an environmental variable
writeRaster(biomass_resamp,"env_variables/Biomass/biomass_WGS84_resamptoprecip.tif", format="GTiff", overwrite = T)
```


# Identify occurrence points
grid at the same resolution as whatever using for the climate data
randomly put down points within polygons 
```{r}
# read in all sugar polygons
all_hi_sugar <- st_read("ag_shapefiles/sugar1900to2020/sugar1900to2020.shp") 

hi_sugar_reproject <- all_hi_sugar %>% 
  st_transform(st_crs(4326)) %>% 
  st_set_precision(1e6) %>% 
  st_make_valid() # class sf and data.frame

```

Need to project the sugarcane shapefile in the same projection and need to get points within the sugarcane shapefile in that projection 

```{r}
s1900 <- st_read("ag_shapefiles/sugar_gee/sugar1900.shp") %>% 
  st_transform(st_crs(4326)) %>% 
  st_set_precision(1e6) %>% 
  st_make_valid()
s1920 <- st_read("ag_shapefiles/sugar_gee/sugar1920.shp") %>% 
  st_transform(st_crs(4326)) %>% 
  st_set_precision(1e6) %>% 
  st_make_valid()
s1937 <- st_read("ag_shapefiles/sugar_gee/sugar1937.shp") %>% 
  st_transform(st_crs(4326)) %>% 
  st_set_precision(1e6) %>% 
  st_make_valid()
ag1980 <- st_read("ag_shapefiles/sugar_gee/ag1980_noblank.shp")%>% 
  st_transform(st_crs(4326)) %>% 
  st_set_precision(1e6) %>% 
  st_make_valid()
ag2015 <- st_read("ag_shapefiles/sugar_gee/ag2015.shp")%>% 
  st_transform(st_crs(4326)) %>% 
  st_set_precision(1e6) %>% 
  st_make_valid()
ag2020 <- st_read("ag_shapefiles/sugar_gee/ag2020.shp")%>% 
  st_transform(st_crs(4326)) %>% 
  st_set_precision(1e6) %>% 
  st_make_valid()

# extract sugar for 2015 and 2020
s2015 <- ag2015 %>% 
  filter(CropCatego == "Sugar")

s2020 <- ag2020 %>% 
  filter(Crops_2020 == "Sugar")
```

```{r}
s1900_1920 <- st_union(s1900, s1920)
s1900_1920_1937 <- st_union(s1900_1920, s1937)
s1900_1920_1937_1980 <- st_union(s1900_1920_1937, s1980)
s1900_1920_1937_1980_2015 <- st_union(s1900_1920_1937_1980, s2015)
s1900_1920_1937_1980_2015_2020 <- st_union(s1900_1920_1937_1980_2015, s2020)
```


# Create reference presence points

Ideally don't want to sample occurrence points randomly. You want to do a tiered sampling of occurrence points so it is 1000 points from fields that were cultivated in 1900, 1000 points from fields cultivated in 1920, ... etc.
```{r}
sugar_centroids <- st_centroid(all_hi_sugar)

sugar_centroids_reproject <- sugar_centroids %>% 
  st_transform(st_crs(4326)) %>% 
  st_set_precision(1e6) %>% 
  st_make_valid() # class sf and data.frame
```
```{r}
sugar_centroid_reproj_coords <- do.call(rbind, st_geometry(sugar_centroids_reproject)) %>% 
  as.tibble() %>% 
  setNames(c("longitude", "latitude")) %>% 
  mutate(name = "sugar")
```

```{r}
sugar_centroid_occ_points <- sugar_centroid_reproj_coords %>% 
  dplyr::select(name, longitude, latitude)
```

```{r}
write.csv(sugar_centroid_occ_points, "occurrences/sugar_centroid_occ_points.csv", row.names = F)
```

```{r}
sugar_points <- st_sample(hi_sugar_reproject, size = 500, type = "random")
```

```{r}
plot(sugar_points)
```



```{r}
sugar_points_coords <- do.call(rbind, st_geometry(sugar_points)) %>% 
  as.tibble() %>% 
  setNames(c("longitude", "latitude")) %>% 
  mutate(name = "sugar")
```

```{r}
sugar_occ_points <- sugar_points_coords %>% 
  dplyr::select(name, longitude, latitude)
```

```{r}
write.csv(sugar_occ_points, "occurrences/sugar_occ_points.csv", row.names = F)
```

```{r}
sugar_points2k <- st_sample(hi_sugar_reproject, size = 2000, type = "random")
```


```{r}
sugar_points2k_coords <- do.call(rbind, st_geometry(sugar_points2k)) %>% 
  as.tibble() %>% 
  setNames(c("longitude", "latitude")) %>% 
  mutate(name = "sugar")
```

```{r}
sugar_occ2k_points <- sugar_points2k_coords %>% 
  dplyr::select(name, longitude, latitude)
```

```{r}
write.csv(sugar_occ2k_points, "occurrences/sugar_occ2k_points.csv", row.names = F)
```

# create the extent from which background points will be drawn
Start with all the land in the Hawaiian Islands
Reduce it to be regions where elevation goes from 0m to the max elevation at which sugarcane is found 

```{r}
hi_background_poly <- st_read("coast_n83/coast_n83.shp") %>% 
  st_transform(st_crs(4326)) %>% 
  st_set_precision(1e6) %>% 
  st_make_valid() # class sf and data.frame
```


```{r}
all_sugar_elev_max <- raster::extract(elev_resamp, hi_sugar_reproject, method = "bilinear")
```

```{r}
sugar_elev_extract = raster::extract(elev_resamp, hi_sugar_reproject)
```

```{r}
elev_poly <- rasterToPolygons(elev_resamp, n = 4, dissolve = F)
```


```{r}
elev_sf <- st_as_sf(elev_poly)
```

```{r}
st_crs(elev_poly)
```



# Run wallace to get the 50% 

```{r}
run_wallace()
```


# visually assessing bare areas
some areas with Ka'u basalt on southern Hawaii Island seem to be bare at 750-1500 years old
also some 1500-3000 year old areas

Puna Basalt 400-750 years ago in vicinity of Volcano
More 400-750 year old flows further north up the coast

South of Kona, Ka'u Basalt 200-750 years ago
Some coastal Hualalai flows from 1500-3000 years ago are bare near Kona

Some bare areas speckled throughout the Hana coast but no single substantial area

doesn't seem to be substantial bare areas on other islands


# Environmental data updates
Ideally you want climate data from the period in which different fields were cultivated (e.g. precipitation from year 1900)

# Background points
How many background points do we want? 10,000?
Constrain background points to land (not ocean)
Constrain background points to elevation range at which sugarcane might be grown
- possibly also try to remove recent volcanic lava flows from the dat as well


# reference field analysis

```{r}
ref_field_run8 <- raster("H:/github/ch1_abandon_patterns/maxent_model_results/Wallace_run8/rm.1_fc.LQ_thresh_p10_run8.tif")
ref_field_run9 <- raster("H:/github/ch1_abandon_patterns/maxent_model_results/Wallace_run9/rm.1_fc.LQ_thresh_p10_run9.tif")
ref_field_run10 <- raster("H:/github/ch1_abandon_patterns/maxent_model_results/Wallace_run10/rm.1_fc.LQ_thresh_p10_run10.tif")
```

```{r}
reclass_matrix = cbind(is = c(0), becomes = c(NA))
```


```{r}
ref_field_run8_reclass <- reclassify(ref_field_run8, reclass_matrix)
ref_field_run9_reclass <- reclassify(ref_field_run9, reclass_matrix)
ref_field_run10_reclass <- reclassify(ref_field_run10, reclass_matrix)
```


